Investigating the Underprovision of Public Goods: Does Segregation Matter?

### 2020-05-03 final_draft_extension_fs.Rmd
changes to city size threshold analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
knitr::write_bib(c("knitr", "stringr"), "bib.bib", width = 60)
library(haven)
library(gtable)
library(stargazer)
library(tidyverse)
library(gt)
library(gtsummary)
library(knitr)
library(stargazer)
library(lme4)
library(margins)
library(reprex)
library(plm)
library(devtools)
library(lfe)
library(Statamarkdown)
library(readxl)
library(starpolishr)
library(AER)
library(readxl)
library(naniar)
library(mice)
library(gtsummary)
library(VIM)
```

@@

rp <- read_dta("racial_polarization_winners.dta")
fin_seg<- read_dta("fin_seg.dta")

# add row numbers for fin_seg

fin_seg$row_num <- as.integer(rownames(fin_seg))

```{r data_cleaning_rp, include = FALSE}

# create a factor for year for the rp dataset
rp$year.f <- as.factor(rp$year)

# apply the condition that winner == 1 as in the Stata code for the rp regressions

rp_1<- rp %>%
  filter(winner == 1) %>%
  mutate(medincinterp = medincinterp/1000)

@@

```{r include = FALSE}

## regression 1 Table 1

# perform a multi-level mixed-effects linear regression of biggest split on the multi-group H index with controls for Herfindahl Diversity Index, % Asian interpolated, % Black interpolated, % Latino interpolated, median household income interpolated, % renters interpolated, percent college grads interpolated, a dummy indicating if there were biracial candidates in the election, a dummy  indicating if the election was nonpartisan, an indicator if the election was a primary, and the log population. The regression also includes fixed effects for year and region and random effects for cities (geo_id2). Diversity is  included because of the claim that more diverse populations will have more heterogeneous political preferences. The other fixed effect control variables are included because they are known to be potentially correlated with segregation. winner == 1 requires that there was a winner in the election.

# regression specification. Apply REML = FALSE. The default for lmer() in R is
# to optimize the restricted (residual) maximum likelihood (REML) whereas in
# xtmixed in Stata the default is to optimize the log likelihood (which you can
# do in lmer() in R by setting REML = FALSE), so this will match the STATA code

m1 <- lmer(biggestsplit ~ H_citytract_multi_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + year.f + south + midwest + west + (1 | geo_id2), data = rp_1, REML=FALSE)

## regression 2 Table 1

# Perform the same regression as for Column 1 above but this time regress on the two-group calculation of Theil's H interpolated. It is unclear where the "diversity" variable is coming from, as the only variable related to diversity in the specificed dataset is diversityinterp. This is a problem that will need to be resolved. Essentially, this regression just tests an alternate specification to the first in which the segregation index in defined in terms of two-groups rather than multiple.

m2 <- lmer(biggestsplit ~ H_citytract_NHW_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + year.f + south + midwest + west + (1 | geo_id2), data = rp_1, REML=FALSE)

# perform the same regression as for column 2, but now add in a measure of
# average white political ideology. The point of this is to control for ideology
# so that the author can separate any effects of segregation from any potential
# effects of white political ideology. This is important because it is possible
# that segregation is just a proxy for white conservatism and thus failure to
# include this variable could lead to ommitted variable bias.

m3 <- lmer(biggestsplit ~ H_citytract_NHW_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + whiteideology_fill2 + year.f + south + midwest + west + (1 | geo_id2), data = rp_1, REML=FALSE)

```{r results="asis", echo = FALSE}

# create a stargazer table of the results

t1 <- stargazer(m1, m2, m3, omit = c("year", "south", "midwest", "west"),
                header = FALSE,
                title = "\\textbf{Racial Polarization in Segregated Cities}",
                single.row = TRUE,
                no.space = FALSE,
                add.lines = list(c("Wald Chi Squarred", "a", "b", "c")),
                omit.stat = c("ll","rsq"),
                covariate.labels = c("Multigroup H Index", "White/Nonwhite H Index", "Diversity", "Percent Asian", "Percent Black", "Percent Latino", "Medain HH Income (1000s)", "Percent Renters", "Percent College Degree", "Biracial Contest", "Nonpartisan Election", "Primary Election", "Population (logged)", "White Ideology", "Constant"),
                notes = "This table displays...")

# note: "This table displays the results of the replication effort of the analyses conducted in Trounstine (2016) using the racial polarization dataset. The replication code can be found on my personal [Github repository](https://github.com/mburzillo/new_final_project). The original table is displayed for comparison as Figure 1.")

![This is the original table displaying the results of the models for the effects of residential segregation on political polarization from page 713 of @trounstine_segregation_2016. My replication of these results can be found in Table 1 in this paper.](table_1.png)

@@

```{r rp_missing_plot, include = FALSE, warning = FALSE}

# don't show the warning, it's just an irrelevant column missing a name

# plot the missing variables if desired

rp_facet <- rp_impute %>%
   select("medincinterp",
             "biggestsplit", "diversityinterp",
             "pctasianpopinterp", "pctblkpopinterp", "pctlatinopopinterp",
             "pctrentersinterp", "pctcollegegradinterp", "biracial", "nonpartisan",
               "south", "midwest", "west", "whiteideology_fill2", "year")

gg_miss_var(rp_impute, show_pct = TRUE) +
  ggtitle("% Missing Values in Trounstine (2016)'s Racial Polarization Dataset")

nrow(rp_impute)

```{r rp_aggr_plot, echo = FALSE}

# select only the missing values

rp_aggr <- rp_impute %>%
   select("medincinterp",
             "biggestsplit", "diversityinterp",
             "pctasianpopinterp", "pctblkpopinterp", "pctlatinopopinterp",
             "pctrentersinterp", "pctcollegegradinterp", "biracial", "nonpartisan",
               "south", "midwest", "west", "whiteideology_fill2")

# create an aggr plot for the missing values

aggr(rp_aggr, bars = TRUE, col=c('navyblue','red'), numbers=TRUE, cex.axis=.5, gap=1, ylab=c("Proporiton of Missing Values","Combinations"))
title(main = "Patterns of Missingness in the Racial Polarization Data",
      sub = "", font.main = 1, font.axis = 1, cex.main = .75)

# caption for this plot when I figure out how to do it... : This plot examines the pattern of missing data in the the original racial polarization dataset from Trounstine (2016). The plot on the left shows the percentage of missing data for each variable with missing data and reveals that the variables that are missing by far the most data are the largest vote split along racial lines (biggestsplit) and the variable indicating whether or not the election had candidates of more than one race (biracial). The plot on the right shows the pattern of missingness among the variables. We can see that approximately 42% of observations are complete (have no missing values). There seems to be a correspondence between missing a value for biggest split and missing biracial. There also seem to be about 2% of values for which most of the variables are missing. However, most observations are not missing more than 2-3 values.

```{r eval = FALSE}
row.plot <- rp_impute %>%
  select(-year.f) %>%
  mutate(id = row_number()) %>%
  gather(-id, key = "key", value = "val") %>%
  mutate(isna = is.na(val)) %>%
  ggplot(aes(key, id, fill = isna)) +
    geom_raster(alpha=0.8) +
    scale_fill_manual(name = "",
        values = c('blue', 'red'),
        labels = c("Present", "Missing")) +
    #scale_x_discrete(limits = levels) +
    labs(x = "Variable",
           y = "Observation Number", 
         title = "Missing values in rows") +
    coord_flip() +
  theme_bw()

row.plot

glimpse(rp_impute)
```

```{r eval = FALSE}

@@

row.plot <- fin_imp %>%
  mutate(id = row_number()) %>%
  gather(-id, key = "key", value = "val") %>%
  mutate(isna = is.na(val)) %>%
  ggplot(aes(key, id, fill = isna)) +
    geom_raster(alpha=0.8) +
    scale_fill_manual(name = "",
        values = c('blue', 'red'),
        labels = c("Present", "Missing")) +
    #scale_x_discrete(limits = levels) +
    labs(x = "Variable",
           y = "Observation Number", 
         title = "Missing values in rows") +
    coord_flip() +
  theme_bw()

row.plot

```{r fin_aggr_plot, echo = FALSE}

fin_imp_aggr <- fin_imp %>%
  select(ideology_fill, H_citytract_NHW_i, chng5pctblk, chng5pctlatino, chng5pctasian,
         diversityinterp, pctblkpopinterp, pctasianpopinterp,
         pctlatinopopinterp, medinc_cpi, pctlocalgovworker_100, pctrentersinterp, pctover65,
         pctcollegegradinterp)

@@

```{r rp_imputation, include = FALSE}

# Perform multiple imputations using the cart method. Because of the time it
# takes to run this function, I turned the results into a dataset to load in
# instead of re-running everytime.

max_it = 20

imp_1_rp <- mice(rp_impute, method = "cart", maxit = 20)

# can use the line below to check that there are now no missing values

# gg_miss_var(complete(imp_1_rp), show_pct = TRUE)

@@

```{r}
#now missing 282261

#before impute:
```

```{r new_imp_less_than_17_nas}

fin_imp <- fin_imp %>%
  mutate(nas = is.na(H_citytract_NHW_i) + is.na(dgepercap_cpi) + is.na(diversityinterp) + is.na(pctblkpopinterp)  + is.na(pctasianpopinterp) + is.na(pctlatinopopinterp) + is.na(medinc_cpi) + is.na(pctlocalgovworker_100) + is.na(pctrentersinterp) + is.na(pctover65) + is.na(pctcollegegradinterp) + is.na(logpop) + is.na(geo_id2) + is.na(chng5pctblk) + is.na(chng5pctlatino) + is.na(chng5pctasian) + is.na(logpop) + is.na(geo_id2) + is.na(chng5pctblk) + is.na(chng5pctlatino) + is.na(chng5pctasian) + is.na(ideology_fill))

fin_imp %>%
  ggplot(aes(x = nas)) +
  geom_histogram()

fin_imp %>%
  group_by(nas) %>%
  summarise(n())

# exclude all observations with 17 out of 22 nas

fin_imp_sub <- fin_imp %>%
  filter(nas < 17)

nrow(fin_imp_sub)

ini <- mice(fin_imp_sub, maxit=0, print=F)

new_imp_fs <- mice(fin_imp_sub, method = meth, maxit = 5)

miss_var_summary(complete(new_imp_fs))

@@

```{r fp_imputations_cities, include = FALSE}

ini <- mice(fin_imp_cities, maxit = 0, print = FALSE)

imp_fs_cities <- mice(fin_imp_cities, method = meth, maxit = 1)

fit_imp_felm1 <- with(imp_fs_cities, felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2))

# RESULTS ARE ESSENTIALLY THE SAME!
print(imp_felm1_sum)

# doesn't add very much to the sample though, still, robust to additional 73 entries

miss_var_summary(complete(imp_fs_cities))

miss_var_summary(fin_imp_cities)

```{r fp_population_full_range}

# what if we condition on population size?

fin_imp_pop_size <- fin_imp %>%
  filter(logpop > 6 & logpop < 16)

nrow(fin_imp_pop_size)

ini <- mice(fin_imp_pop_size, maxit = 0, print = FALSE)

imp_fs_pop_size <- mice(fin_imp_pop_size, method = meth, maxit = 5)

fit_imp_felm1 <- with(imp_fs_pop_size, felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2))

# RESULTS ARE ESSENTIALLY THE SAME!
print(imp_felm1_sum)

# doesn't add very much to the sample though, still, robust to additional 73 entries

miss_var_summary(complete(imp_fs_cities))

miss_var_summary(fin_imp_cities)

```{r 95_pop}
fin_imp_pop_trounstine_95 <- fin_imp %>%
  filter(logpop < 11.93437 & logpop > 8.7754)

nrow(fin_imp_pop_trounstine_95)

ini <- mice(fin_imp_pop_trounstine_95, maxit = 0, print = FALSE)

imp_fs_pop_size_trounstine_95 <- mice(fin_imp_pop_trounstine_95, method = meth, maxit = 5)

fit_imp_felm1 <- with(imp_fs_pop_size_trounstine_95, felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2))

# RESULTS ARE ESSENTIALLY THE SAME!
print(imp_felm1_sum)

# doesn't add very much to the sample though, still, robust to additional 73 entries

miss_var_summary(complete(imp_fs_cities))

miss_var_summary(fin_imp_cities)

```{r population_first_quartile}

quantile(fin_dge$logpop, c(.25))

quantile(fin_seg$logpop, c(.25), na.rm = T)

fin_imp_pop_trounstine_first_quart <- fin_imp %>%
  filter(logpop < 9.388152)

nrow(fin_imp_pop_trounstine_first_quart)

gg_miss_var(fin_imp_pop_trounstine_first_quart, show_pct = T)

ini <- mice(fin_imp_pop_trounstine_first_quart, maxit = 0, print = FALSE)

imp_fs_pop_size_bottom_25 <- mice(fin_imp_pop_trounstine_first_quart, method = meth, maxit = 5)

fit_imp_felm1 <- with(imp_fs_pop_size_bottom_25, felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2))

# RESULTS ARE ESSENTIALLY THE SAME!
print(imp_felm1_sum)

# doesn't add very much to the sample though, still, robust to additional 73 entries

miss_var_summary(complete(imp_fs_cities))

miss_var_summary(fin_imp_cities)

```{r population_top_90_trounstine}

quantile(fin_dge$logpop, c(.90))

quantile(fin_seg$logpop, c(.90), na.rm = T)

fin_imp_pop_top_90_trounstine <- fin_imp %>%
  filter(logpop > 11.43575)

fin_seg %>%
  ggplot() +
  geom_histogram(aes(x = logpop))

nrow(fin_imp_pop_top_90_trounstine)

gg_miss_var(fin_imp_pop_top_90_trounstine, show_pct = T)

ini <- mice(fin_imp_pop_top_90_trounstine, maxit = 0, print = FALSE)

imp_fs_pop_size_top_90_trounstine <- mice(fin_imp_pop_top_90_trounstine, method = meth, maxit = 5)

fit_imp_felm1 <- with(imp_fs_pop_size_top_90_trounstine, felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2))

# RESULTS ARE ESSENTIALLY THE SAME!
print(imp_felm1_sum)

# doesn't add very much to the sample though, still, robust to additional 73 entries

miss_var_summary(complete(imp_fs_cities))

miss_var_summary(fin_imp_cities)

```{r population_bottom_5_trounstine}

quantile(fin_dge$logpop, c(.05))

quantile(fin_seg$logpop, c(.05), na.rm = T)

fin_imp_pop_bottom_5_trounstine <- fin_imp %>%
  filter(logpop < 8.775581)

fin_seg %>%
  ggplot() +
  geom_histogram(aes(x = logpop))

nrow(fin_imp_pop_bottom_5_trounstine)

gg_miss_var(fin_imp_pop_bottom_5_trounstine, show_pct = T)

ini <- mice(fin_imp_pop_bottom_5_trounstine, maxit = 0, print = FALSE)

imp_fs_pop_size_bottom_5_trounstine <- mice(fin_imp_pop_bottom_5_trounstine, method = meth, maxit = 1)

fit_imp_felm1 <- with(imp_fs_pop_size_bottom_5_trounstine, felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2))

# RESULTS ARE ESSENTIALLY THE SAME!
print(imp_felm1_sum)

# doesn't add very much to the sample though, still, robust to additional 73 entries

miss_var_summary(complete(imp_fs_pop_size_bottom_5_trounstine))

miss_var_summary(fin_imp_cities)

nrow(complete(imp_fs_pop_size_bottom_5_trounstine))

```{r population_bottom_1}

quantile(fin_dge$logpop, c(.01))

quantile(fin_seg$logpop, c(.20), na.rm = T)

fin_imp_pop_bottom_1_trounstine <- fin_imp %>%
  filter(logpop < 8.364976 )

fin_seg %>%
  ggplot() +
  geom_histogram(aes(x = logpop))

nrow(fin_imp_pop_bottom_1_trounstine)

gg_miss_var(fin_imp_pop_bottom_1_trounstine, show_pct = T)

ini <- mice(fin_imp_pop_bottom_1_trounstine, maxit = 0, print = FALSE)

imp_fs_pop_size_bottom_1_trounstine <- mice(fin_imp_pop_bottom_1_trounstine, method = meth, maxit = 1)

fit_imp_felm1 <- with(imp_fs_pop_size_bottom_1_trounstine, felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2))

# RESULTS ARE ESSENTIALLY THE SAME!
print(imp_felm1_sum)

# doesn't add very much to the sample though, still, robust to additional 73 entries

miss_var_summary(complete(imp_fs_cities))

miss_var_summary(fin_imp_cities)

```{r population_over_Baqir}

fit_imp_felm1 <- with(imp_fs_pop_overbaqir, felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2))

# RESULTS ARE ESSENTIALLY THE SAME!
print(imp_felm1_sum)

# doesn't add very much to the sample though, still, robust to additional 73 entries

miss_var_summary(complete(imp_fs_cities))

miss_var_summary(fin_imp_cities)

```{r population_under_Baqir}

fit_imp_felm1 <- with(imp_fs_pop_under_baqir, felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2))

# RESULTS ARE ESSENTIALLY THE SAME!
print(imp_felm1_sum)

# doesn't add very much to the sample though, still, robust to additional 73 entries

miss_var_summary(complete(imp_fs_cities))

miss_var_summary(fin_imp_cities)

```{r population_mid_50}

# essentially the same - let's do full population
(quantile(fin_seg$logpop, c(.25, .75), na.rm = T))
(quantile(fin_imp$logpop, c(.25, .75), na.rm = T))

fin_imp_pop_mid_50 <- fin_imp %>%
  filter(logpop > 5.777652 & logpop < 7.936303)

nrow(fin_imp_pop_mid_50)

gg_miss_var(fin_imp_pop_mid_50, show_pct = T)

ini <- mice(fin_imp_pop_mid_50, maxit = 0, print = FALSE)

imp_fs_pop_mid_50 <- mice(fin_imp_pop_mid_50, method = meth, maxit = 1)

fit_imp_felm1 <- with(imp_fs_pop_mid_50, felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2))

# RESULTS ARE ESSENTIALLY THE SAME!
print(imp_felm1_sum)

# doesn't add very much to the sample though, still, robust to additional 73 entries

miss_var_summary(complete(imp_fs_cities))

miss_var_summary(fin_imp_cities)

```{r population_top_1_trounstine}

(quantile(fin_seg$logpop, c(.99), na.rm = T))
(quantile(fin_dge$logpop, c(.99), na.rm = T))

fin_imp_top_1_trounstine <- fin_imp %>%
  filter(logpop > 13.1907 )

nrow(fin_imp_top_1_trounstine)

gg_miss_var(fin_imp_top_1_trounstine, show_pct = T)

ini <- mice(fin_imp_top_1_trounstine, maxit = 0, print = FALSE)

imp_fs_pop_top_1_trounstine <- mice(fin_imp_top_1_trounstine, method = meth, maxit = 5)

fit_imp_felm1 <- with(imp_fs_pop_top_1_trounstine, felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2))

# RESULTS ARE ESSENTIALLY THE SAME!
print(imp_felm1_sum)

nrow(complete(imp_fs_pop_top_1_trounstine))

# doesn't add very much to the sample though, still, robust to additional 73 entries

miss_var_summary(complete(imp_fs_cities))

miss_var_summary(fin_imp_cities)

```{r population_top_1}

(quantile(fin_seg$logpop, c(.99), na.rm = T))
(quantile(fin_dge$logpop, c(.99), na.rm = T))

fin_imp_top_1 <- fin_imp %>%
  filter(logpop > 11.28648)

nrow(fin_imp_top_1)

gg_miss_var(fin_imp_top_1, show_pct = T)

ini <- mice(fin_imp_top_1, maxit = 0, print = FALSE)

imp_fs_pop_top_1 <- mice(fin_imp_top_1_trounstine, method = meth, maxit = 5)

fit_imp_felm1 <- with(imp_fs_pop_top_1, felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2))

# RESULTS ARE ESSENTIALLY THE SAME!
print(imp_felm1_sum)

nrow(complete(imp_fs_pop_top_1))

# doesn't add very much to the sample though, still, robust to additional 73 entries

miss_var_summary(complete(imp_fs_cities))

miss_var_summary(fin_imp_cities)

Before examining the results of Trounstine's model using the multiply imputed data, I first ran some diagnostic tests on the imputation results to make sure that everything ran as expected. First, I checked the convergence of the algorithm used within mice() by plotting the trace lines as a function of the number of iterations for each of the variables. Then, I visually inspected the distributions of the imputed data in comparison to the original data with density and strip plots. All of these checks suggested that the imputed values were within a plausible range of the data and that their distribution fit the underlying distribution of the data relatively well. The only cases in which there was some cause for concern were for some of the variables with very few missing values in the racial polarization dataset (such as the indicators for region). However, this is more or less to be expected given the small number of imputations performed in these cases. Thus, and especially because there are so few of these values missing in the actual datasets, this was not a major concern. See figures X and X1 to see the density plots of the imputed values overlayed on the density plots of the original variables. The rest of the plots and results of the diagnostic tests discussed here are presented in the Appendix.

```{r rp_densityplot, echo = FALSE}

densityplot(imp_1_rp)

```{r rp_imp_plot, include = FALSE}

# ideally, the fits will intertwine and not exhibit any trends at later iterations. This seems to generally be the case. There are some issues with the results for some of the variables with very few missing values, which makes sense as we would expect the mean to be less reliable due to the law of large numbers and thus convergence is more difficult. However, since these are so few values, it is less important about how robust they are.

plot(imp_1_rp)

```{r rp_stripplot, include = FALSE}

# we can also check the imputed values against the original values using stripplot(). Each column in each subplot represents a separate iteration. The magenta points represent the imputed data. The values of the variable in questions are along the y axis. We expect the spread of the data to be similar if the imputations were done well. If the data were missing completely at random, then the imputed data should have the same distribution as the original data. In particular, we want to be sure that the imputations are within a plausible range of the data. This is the case for all of our imputed variables, and there does not seem to be cause for alarm from these results.

# this looks good? All reasonable values...

stripplot(imp_1_rp)

stripplot(imp_1_rp, biggestsplit)

```{r fs_densityplot, echo = FALSE}

densityplot(imp_fs)
densityplot(x=imp_fs , data= ~ H_citytract_NHW_i + chng5pctblk + 
       chng5pctlatino + chng5pctasian + ideology_fill)

```{r fs_imp_plot, include = FALSE}

# ideally, the fits will intertwine and not exhibit any trends at later iterations. This seems to generally be the case. There are some issues with the results for some of the variables with very few missing values, which makes sense as we would expect the mean to be less reliable due to the law of large numbers and thus convergence is more difficult. However, since these are so few values, it is less important about how robust they are.

plot(imp_fs)

```{r fs_stripplot, include = FALSE}

# we can also check the imputed values against the original values using stripplot(). Each column in each subplot represents a separate iteration. The magenta points represent the imputed data. The values of the variable in questions are along the y axis. We expect the spread of the data to be similar if the imputations were done well. If the data were missing completely at random, then the imputed data should have the same distribution as the original data. In particular, we want to be sure that the imputations are within a plausible range of the data. This is the case for all of our imputed variables, and there does not seem to be cause for alarm from these results.

# this looks good? All reasonable values...

stripplot(imp_fs, H_citytract_NHW_i)

stripplot(imp_fs, chng5pctblk)

stripplot(imp_fs, chng5pctlatino)

stripplot(imp_fs, chng5pctasian)

stripplot(imp_fs, ideology_fill)

Given the promising results of the diagnostic checks, I next proceeded to re-estimate the original model using the multiply imputed datasets, pooling the results to produce final pooled regression coefficients and parameters. The results for the analyses using the racial polarization dataset are presented in tables 1-3.

```{r rp_imputations, include = FALSE}

# Now that we have imputed the missing data, we can re-run the analyses from the
# original paper to see how and if this changes the analysis. We can use the
# with() function from the mice package to fit each of the imputed datasets to
# the model and then pool the results afterwards.

# regression 1: fit multiple imputed datasets

fit_imp_1_rp <- with(imp_1_rp, lmer(formula = biggestsplit ~ H_citytract_multi_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + year.f + south + midwest + west + (1 | geo_id2), REML=FALSE))

# pool and summarize the analyses

pool_imp_1_rp <- pool(fit_imp_1_rp)
imp_1_rp_sum <- summary(pool_imp_1_rp)

## regression 2 Table 1: fit multiple imputed datasets

#formula = biggestsplit ~ H_citytract_NHW_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + year.f + south + midwest + west + (1 | geo_id2)

fit_imp_2_rp <- with(imp_1_rp, lmer(biggestsplit ~ H_citytract_NHW_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + year.f + south + midwest + west + (1 | geo_id2), REML=FALSE))

# pool and summarize the analyses

pool_imp_2_rp <- pool(fit_imp_2_rp)
imp_2_rp_sum <- summary(pool_imp_2_rp)

## regression 3 Table 1 with new data

#formula = biggestsplit ~ H_citytract_NHW_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + whiteideology_fill2 + year.f + south + midwest + west + (1 | geo_id2)

fit_imp_3_rp <- with(imp_1_rp, lmer(biggestsplit ~ H_citytract_NHW_i + diversityinterp + pctasianpopinterp + pctblkpopinterp + pctlatinopopinterp + medincinterp + pctrentersinterp +  pctcollegegradinterp + biracial + nonpartisan + primary + logpop + whiteideology_fill2 + year.f + south + midwest + west + (1 | geo_id2), REML=FALSE))

# pool and summarize the analyses

pool_imp_3_rp <- pool(fit_imp_3_rp)
imp_3_rp_sum <- summary(pool_imp_3_rp)

@@

```{r results_printing_rp_impute_regressions_1, results = "asis", echo=FALSE}

imp_1_rp_sum_table <- as.data.frame(imp_1_rp_sum %>% 
  select(term, estimate, std.error, p.value)) %>%
  head(13)
print(imp_1_rp_sum_table)

stargazer(imp_1_rp_sum_table,
          header = FALSE,
          title = "\\textbf{Racial Polarization in Segregated Cities: Racial Divide with Multigroup Segregation Index}")

In the second and third regressions, the results are similar. The main coefficient on segregation index has decreased in magnitude, although in these specifications, it remains statistically significant. The other coefficients except for the biracial indicator are insignificant, and the coefficients on percent Asian, Black, and Latino have all switched signs. Standard errors have also reduced slightly in these specifications compared to the original analysis in @trounstine_segregation_2016.

```{r results_printing_rp_impute_regressions_2, results = "asis", echo=FALSE}

imp_2_rp_sum_table <- as.data.frame(imp_2_rp_sum %>% 
  select(term, estimate, std.error, p.value)) %>%
  head(13)
print(imp_2_rp_sum_table)

stargazer(imp_2_rp_sum_table,
          header = FALSE,
          title = "\\textbf{Racial Polarization in Segregated Cities: Racial Divide with Two-Group Segregation Index}")

```{r results_printing_rp_impute_regressions_3, results = "asis", echo=FALSE}

>>>>

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
knitr::write_bib(c("knitr", "stringr"), "bib.bib", width = 60)
library(haven)
library(stargazer)
library(tidyverse)
library(gtsummary)
library(knitr)
library(AER)
library(lme4)
library(margins)
library(lfe)
library(readxl)
library(mice)
library(VIM)
library(kableExtra)
library(huxtable)
```

@@

#rp <- read_dta("racial_polarization_winners.dta")
fin_seg<- read_dta("fin_seg.dta")

@@

@@

```{r fin_aggr_plot, echo = FALSE, fig.cap = "This plot examines the pattern of missing data in the the original financial segregation dataset from Trounstine (2016). The plot on the left shows the percentage of missing data for each variable with missing data and reveals that a large number of variables are missing the vast majority of data. The plot on the right shows the pattern of missingness among the variables. We can see that only .58\\% of observations are complete (have no missing values), and a large majority of the observations are missing data for many variables.", fig.width= 7, fig.height = 5, fig.pos="H", eval = FALSE}

@@

fin_imp_aggr <- fin_imp %>%
  select(ideology_fill, H_citytract_NHW_i, chng5pctblk, chng5pctlatino, 
         chng5pctasian, diversityinterp, pctblkpopinterp, pctasianpopinterp,
         pctlatinopopinterp, medinc_cpi, pctlocalgovworker_100, pctrentersinterp,
         pctover65, pctcollegegradinterp) %>%
    rename("White Ideology" = ideology_fill, "White/Non-White H Index" = H_citytract_NHW_i,
           "5Y Change % Black" = chng5pctblk, "5Y Change % Latino" = chng5pctlatino, 
           "5Y Change % Asian" = chng5pctasian, "Diversity" = diversityinterp, 
           "% Asian" = pctasianpopinterp, "% Black" = pctblkpopinterp, 
           "% Latino" = pctlatinopopinterp, "Median HH Income" = medinc_cpi,
           "% Local Gov. Worker" = pctlocalgovworker_100, "% Renters" = pctrentersinterp, 
           "% College Grads" = pctcollegegradinterp, "% 65+" = pctover65)

@@

@@

Before examining the results of Trounstine's model using the multiply imputed data, I first ran some diagnostic tests on the imputation results to make sure that everything ran as expected. First, I checked the convergence of the algorithm used within mice() by plotting the trace lines as a function of the number of iterations for each of the variables. Then, I visually inspected the distributions of the imputed data in comparison to the original data with density and strip plots. All of these checks suggested that the imputed values were within a plausible range of the data and that their distribution fit the underlying distribution of the data relatively well. The only cases in which there was some cause for concern were for some of the variables with very few missing values in the racial polarization dataset (such as the indicators for region). However, this is more or less to be expected given the small number of imputations performed in these cases. Thus, and especially because there are so few of these values missing in the actual datasets, this was not a major concern. See figures X and X1 to see the density plots of the imputed values overlayed on the density plots of the original variables. The code to generate the rest of the plots and results of the diagnostic tests discussed here are presented in the Appendix.

```{r fs_densityplot, echo = FALSE, eval = FALSE}

my.settings <- list(
  par.main.text = list(font = 1, cex = 1)
)

#densityplot(imp_fs)
densityplot(x = imp_fs , 
            data = ~ H_citytract_NHW_i + chng5pctblk + chng5pctlatino + chng5pctasian + ideology_fill,
            main = "Density Plots of Imputed and Original Data by Variable\nin the Financial Segregation Dataset",
            par.settings = my.settings,
            par.strip.text = list(cex = 1))

@@

Given the promising results of the diagnostic checks, I next proceeded to re-estimate the original model using the multiply imputed datasets, pooling the results to produce final pooled regression coefficients and parameters. The results for the analyses using the racial polarization dataset are presented in tables X1-X3.

@@

In the second regression, the results are similar. The main coefficient on segregation index has decreased in magnitude, although in these specifications, it remains statistically significant. The other coefficients except for the biracial indicator are insignificant, and the coefficients on percent Black, percent Latino, and median household income have all switched signs.

In the third regression, like in the second, the main coefficient on segregation index has decreased in magnitude but remains significant. The other coefficients except for the biracial indicator are insignificant, and the coefficients on percent Black, percent Asian, median income, and logged population have switched signs. In all three of the specifications, tandard errors reduced slightly using the larger, multiply imputed datasets compared to the original analysis with the original data in @trounstine_segregation_2016.
